### 2024-07-26 modern-portfolio-theory.qmd
20240725 - cluster analysis
20240725 - updates

Modern portfolio theory is a framework for determining the optimal composition of an investment portfolio to maximize expected returns for a given level of risk.
Here, *risk* refers to the [variability](#sec-descriptiveStatisticsSpread) (e.g., standard deviation or variance) of returns across time.
Given two portfolios with the same expected returns over time, people will prefer the "safer" portfolio—that is, the portfolio with less [variability](#sec-descriptiveStatisticsSpread)/volatility across time.
One of the powerful notions of modern portfolio theory is that, through diversification, one can achieve lower risk with the same expected returns.
In investing, *diversification* involves owning multiple asset classes (e.g., domestic and international stocks and bonds), with the goal of having asset classes that are either uncorrelated or negatively correlated.
That is, owning different types of assets is safer than owning only one type.
If you have too much money in one asset and that asset tanks, you will lose your money.
In other words, you do not want to put all of your eggs in one basket.
By owning different asset classes, you can limit your downside risk without sacrificing much in terms of expected return.

@@

Another important idea from modern portfolio theory is that, if you want to achieve higher returns, you may be able to by accepting additional—and the right combination of—risk.
That is, higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.

@@

The ultimate goal in fantasy football is to draft players for your starting lineup that provide the most projected points and the smallest downside risk.
That is, your goal is to achieve the optimal portfolio at a given level of risk, depending on how much risk you are willing to tolerate.
One of the key tools in modern portfolio theory for identifying the optimal portfolio (for a given risk level) is the efficient frontier.
The efficient frontier is a visual depiction of the maximum expected returns for a given level of risk (where risk is the [variability](#sec-descriptiveStatisticsSpread) in returns over time).
The efficient frontier is helpful for identifying the optimal portfolio—the optimal combination and weighting of assets—for a given risk level.

@@

>>>>

Modern portfolio theory (mean-variance theory) is a framework for determining the optimal composition of an investment portfolio to maximize expected returns for a given level of risk.
Here, *risk* refers to the [variability](#sec-descriptiveStatisticsSpread) (e.g., standard deviation or variance) of returns across time.
Given two portfolios with the same expected returns over time, people will prefer the "safer" portfolio—that is, the portfolio with less [variability](#sec-descriptiveStatisticsSpread)/volatility across time.
One of the powerful notions of modern portfolio theory is that, through diversification, one can achieve lower risk with the same expected returns.
In investing, *diversification* involves owning multiple asset classes (e.g., domestic and international stocks and bonds), with the goal of having asset classes that are either uncorrelated or negatively correlated.
That is, owning different types of assets is safer than owning only one type.
If you have too much money in one asset and that asset tanks, you will lose your money.
In other words, you do not want to put all of your eggs in one basket.
By owning different asset classes, you can limit your downside risk without sacrificing much in terms of expected return.

@@

Another important idea from modern portfolio theory is that, if you want to achieve higher returns, you may be able to by accepting additional—and the right combination of—risk.
In general, risk is positively correlated with return.
That is, receiving higher returns generally requires taking on additional risk—at least as long as we stay along the [efficient frontier](#sec-efficientFrontier), described next.

@@

The ultimate goal in fantasy football is to draft players for your starting lineup that provide the most projected points and the smallest downside risk.
That is, your goal is to achieve the optimal portfolio at a given level of risk, depending on how much risk you are willing to tolerate.
One of the key tools in modern portfolio theory for identifying the optimal portfolio (for a given risk level) is the efficient frontier.
The efficient frontier is a visual depiction of the maximum expected returns for a given level of risk (where risk is the [variability](#sec-descriptiveStatisticsSpread) in returns over time).
The efficient frontier is helpful for identifying the optimal portfolio—the optimal combination and weighting of assets—for a given risk level.
Anything below the efficient frontier is considered inefficient (i.e., lower-than-maximum returns for a given level of risk).

@@

### Based on Variability Across Projection Sources {#sec-efficientFrontierFantasyProjectionsUncertainty}

### Based on Historical Game-to-Game Variability {#sec-efficientFrontierFantasyProjectionsUncertainty}

### 2024-07-26 sports-cognitive-psychology.qmd
20240725 - updates

>>>>

## Overview {#sec-sportsCognitivePsychologyOverview}

<https://stacker.com/sports/15-ways-analytics-has-changed-sports> (archived at <https://perma.cc/PQ5R-TWFA>)

- Bill James, "Moneyball", Billy Beane, Theo Epstein, and role of Sabermetrics in helping MLB teams with limited budgets (e.g., Oakland Athletics) compete with larger budgets: using statistics to identify player value more accurately, especially for identifying undervalued players
- development of advanced metrics (e.g., value over replacement player, wins above replacement, fieling independent pitching)
- focus on on-base percentage and slugging percentage (and on-base plus slugging percentage) instead of batting average; defensive shifting; fewer attempts to steal bases; fewer bunts; batters taking more pitches; more frequent pitching changes (for pitcher–batter matchups, e.g., righty–lefty); focus on velocity and spin rate
- other sports, including basketball and football are making more use of analytics
- basketball: types of shots (focusing on three pointers and layups); star players resting more games
- football: going for it on fourth down; greater emphasis on the passing game; drafting Running Backs later in the draft (and, overall, valuing Running Backs less); trading down in the draft for more low picks than fewer high picks [because top picks are frequently overvalued; @Massey2013]
